% alt + shift + 1
% make projname.pdf
% wordcount: detex writeup.tex | wc -w
\documentclass[defaultstyle,11pt]{thesis}
%------------------------Dimensions----------------------------
\topmargin=0.0in
\oddsidemargin=0.0in           % 1in margins at left and right
\evensidemargin=0in
\textwidth=6.5in               % US paper is 8.5in wide
\marginparwidth=0.5in

\headheight=0pt                % 1in margins at top and bottom
\headsep=0pt
\topmargin=0in
\textheight=9.0in              % US paper is 11.0in high

%adjustments...
\addtolength{\topmargin}{-0.5in}
\addtolength{\textheight}{1.0in}
\addtolength{\textwidth}{0.5in}
\addtolength{\oddsidemargin}{-0.25in}
\addtolength{\evensidemargin}{-0.25in}                       
%------------------------Packages----------------------------
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage[paperwidth=8.5in,paperheight=11in,margin=0.98in]{geometry}
\usepackage{listings,float}
\usepackage{color}
\usepackage{array}
\usepackage{cancel}
\usepackage[numbers]{natbib}
\captionsetup{width=.75\textwidth} 
%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Spatially Random Processes in One-Dimensional Maps: The
  Logistic Map and The Arnold Circle Map}

\author{A.~T.}{Le}

\otherdegrees{B.S., University of Colorado, 2015}

\degree{Master of Science}		%  #1 {long descr.}
	{M.S., Applied Mathematics}		%  #2 {short descr.}

\dept{Department of}			%  #1 {designation}
	{Applied Mathematics}

\advisor{}				%  #1 {title}
	{Dr.~James Meiss}			%  #2 {name}

\reader{Dr.~Juan Restrepo}		%  2nd person to sign thesis
\readerThree{Dr.~William Kleiber}		%  3rd person to sign thesis

\abstract{  \OnePageChapter	% because it is very short
One way to model in-situ remediation of contaminated groundwater is to
consider spatially random processes in nonlinear systems. Groundwater remediation often
requires injecting an aquifer with treatment solution, where
degredation reactions break down the toxins. As the treatment solution
and contaminated water flow through the aquifer, their movement is
limited by the types of sediment found in the aquifer, which act as
spatial barriers to mixing. The onset of chaos in
this system implies the two solutions are well mixed, and thus the
contaminants are rendered inert. The spatially random processes
explored in this thesis are meant to mimic the distribution of
sediment in the aquifer. These processes were constructed using
uniform random variables and normal random variables, and incorporate
an exponentially decaying spatial correlation. 

The three-dimensional model of the fluid flow
in the aquifer has been simplified to an in-deptn study of two
one-dimensional maps: the logistic map and the Arnold circle
map. Injection of the treatment solution in the aquifer may be thought
of as the initial condition imposed on the map. Numerical simulations
of the one-dimensional maps lay the groundwork for future studies of
higher-dimensional systems. 

Simulations indicate evidence of newly stabilized regions of the
randomized logistic map, as well as a breakdown of symmetry and stable
behavior in the Arnold circle map. The combination of bifurcation
diagrams and Lyapunov exponents from the randomized logistic map lead
us to hypothesize the spatially random process may stabilize the map
in regions previously unstable. In the random circle map, analysis of
the Arnold tongues, devil's staircases, and Lyapunov exponents suggest
the random processes incur chaotic behavior in typically stable regions.
	}

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To my mom, who has always told me, ``Just do your best.''
	}

\acknowledgements{	\OnePageChapter	% *MUST* BE ONLY ONE PAGE!
For his extensive patience, support, and guidance, I am truly grateful to Jim
Meiss. For being on my committee, I offer thanks to Jim Meiss, Juan Restrepo,
and Will Kleiber. For his unwavering confidence in me and continued
encouragement, I have sincere gratitude for Swamy Ananthanarayan. 
	}

% \IRBprotocol{E927F29.001X}	% optional!

\ToCisShort	% use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input macros.tex
\input chapter1.tex
\input chapter2.tex
\input chapter3.tex
\input chapter4.tex
%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain}	% or "siam", or "alpha", etc.
\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input appendixA.tex
\input appendixB.tex

\end{document}

